Matching the Time of Emergence of Transboundary Fish Stocks to Lead Time for

### 2020-06-15 Scripts/Initial_Analysis.Rmd
Fixed code and made first result run

  # ----------------- #
  # Get model data from spps for one ensemble
  #  ---------------- #

@@

  #### Creates a data to control computational time (lol!)
    if(file.exists(paste(my_path("R"),"spp_df.csv",sep="")) == FALSE){
    spp_df <- tibble() %>% 
      write_csv(.,
                paste(my_path("R"),"spp_df.csv",sep=""))
  }

  suppressMessages(
    spp_df <- my_path("R",name="spp_df.csv",read = TRUE)
  )
  spp_df <- tibble(taxon_key,
                   ensemble) %>% 
    bind_rows(spp_df) %>% 
    write_csv(.,
              paste(my_path("R"),"spp_df.csv",sep=""))

@@

# GetCentroids(taxon_key = 600006,
#              year = year_df$year[1:5],
#              data_type = "Catch",
#              ensemble = 102,
#              Neighbours = Neighbours,
#              eez_centroid = eez_centroid
#              )

@@

@@

  # Historical mean and sd of distance
  historical <- Centroids_id %>%
    filter(year <= Hist_y) %>% 
    ungroup() %>%
    group_by(taxon_key,ensemble,eez_name,neighbour_id) %>%
    summarise_at(vars("distance"),
                 c(hd_temp_mean=mean,
                   hd_temp_sd=sd)
    )

@@

  # Transboundary Index
  Trans_index <- Centroids_id %>% 
    left_join(historical,
              by = c("taxon_key","eez_name","ensemble","neighbour_id")
    ) %>%
    # Re arrenging table for mutate
    gather("variable","value",distance:hd_temp_sd) %>%
    ungroup()

  # Get unique eezs
  Unique_EEZs <- unique(Trans_index$eez_name)

  Partial <- Trans_index %>% 
    mutate(
      names = ifelse(eez_name ==Unique_EEZs[1],paste("Country_A",variable,sep="_"),paste("Country_B",variable,sep="_"))
    ) %>%
    select(-eez_name,-variable) %>%
    spread(names,value) %>% 
    # mutate index
    mutate(
      trans_index = (Country_A_distance/Country_A_hd_temp_sd - Country_B_distance/Country_B_hd_temp_sd)^2
    ) %>% 
    # Ten years average of results
    group_by(taxon_key,ensemble,neighbour_id) %>% 
    mutate(RMean = rollmean(x = trans_index, 
                            10, 
                            align = "right", 
                            fill = trans_index)
    ) %>% 
    filter(year > 1960) %>% 
    select(taxon_key,neighbour_id,year,ensemble,trans_index,RMean)

  # Estimate noice from ensemble members
  historic_sd <- Partial %>% 
    filter(year < Hist_y) %>% 
    group_by(taxon_key) %>% 
    summarise_at(vars("RMean"),
                 c(hist_mean=mean,
                   noise=sd)
    )

@@

  # Identifies the year of emergence
  Final_Result <- mean_sd_index %>%
    filter(year > Hist_y,
           trend > (historic_sd$hist_mean+historic_sd$noise) |
             trend < (historic_sd$hist_mean-historic_sd$noise)
    ) %>%
    group_by(taxon_key,neighbour_id) %>%
    summarise(emerging_yr= min(year)) %>%
    left_join(EEZ_ids) %>%
    ungroup() %>%
    select(taxon_key,eez_name,eez_neighbour,emerging_yr)

  return(Final_Result)
}

@@

GetTransIndex <- function(taxon_key,
                          year,
                          Hist_y,
                          data_type = "Catch",
                          path,
                          ensemble_list,
                          Neighbours,
                          eez_centroid){

    # First step
    # Call Get centroids fucntion for centroids data per ensemble
    Step_one <- bind_rows(
      lapply(ensemble_list[1:2],
             GetCentroids,
             taxon_key = taxon_key,
             year = year_df$year,
             data_type = data_type,
             Neighbours = Neighbours,
             eez_centroid = eez_centroid
             )
    )

@@

@@

    return(Step_two)
  }

  # GetTransIndex(
  #   taxon_key = 600006,
  #   ensemble_list = ensemble_list,
  #   year = year_df$year,
  #   Hist_y = 2000,
  #   data_type = "Abd",
  #   Neighbours = Neighbours,
  #   eez_centroid = eez_centroid
  # )

@@

#  List os species
spp_list <- unique(Transboundary_spp$taxon_key)[1:2]

@@

####_______________________ ####
# Estimating year of eergence
####_______________________ ####

suppressMessages(
  suppressWarnings(
    Emerging_result <- bind_rows(
      mclapply(spp_list[1],
               GetTransIndex,
               year = year_df$year,
               Hist_y = Historic_year,
               data_type = "Catch",
               ensemble_list = ensemble_list,
               # path = path,
               Neighbours = Neighbours,
               eez_centroid = eez_centroid
      )
    )
  )
)

>>>>

  # ----------------- #
  # Get model data from spps for one ensemble
  #  ---------------- #
 if(ensemble == 102){
   print(taxon_key)
 }

@@

  #### Creates a data to control computational time (lol!)
  #   if(file.exists(paste(my_path("R"),"spp_df.csv",sep="")) == FALSE){
  #   spp_df <- tibble() %>% 
  #     write_csv(.,
  #               paste(my_path("R"),"spp_df.csv",sep=""))
  # }
  # 
  # suppressMessages(
  #   spp_df <- my_path("R",name="spp_df.csv",read = TRUE)
  # )
  # spp_df <- tibble(taxon_key,
  #                  ensemble) %>% 
  #   bind_rows(spp_df) %>% 
  #   write_csv(.,
  #             paste(my_path("R"),"spp_df.csv",sep=""))

@@

# GetCentroids(taxon_key = 600076,
#              year = year_df$year[1:5],
#              data_type = "Catch",
#              ensemble = 102,
#              Neighbours = Neighbours,
#              eez_centroid = eez_centroid
#              )

@@

Note: If no historic data available then emerging year is the first year of data in an EEZ

@@

  # Historical mean and sd of distance
  historical <- Centroids_id %>%
    filter(year <= Hist_y) %>% 
    ungroup() %>%
    group_by(taxon_key,ensemble,eez_name,neighbour_id) %>%
    summarise_at(vars("distance"),
                 c(hd_temp_mean=mean,
                   hd_temp_sd=sd,
                   n_yrs =~n())
                 )

@@

   # Get unique eezs
  Unique_EEZs <- unique(EEZ_ids$eez_name)

  # Transboundary Index
  Trans_index <- Centroids_id %>% 
    left_join(historical,
              by = c("taxon_key","eez_name","ensemble","neighbour_id")
    ) %>%
    # Re arrenging table for mutate
    gather("variable","value",distance:hd_temp_sd) %>%
    ungroup() %>% 
    mutate(value = ifelse(is.na(value),0,value)) %>% 
    mutate(
      names = ifelse(eez_name ==Unique_EEZs[1],paste("Country_A",variable,sep="_"),paste("Country_B",variable,sep="_")
    )
    ) %>%
    select(-eez_name,-variable) %>%
    spread(names,value) %>% 
    # mutate index
    mutate(
      trans_index = (Country_A_distance/Country_A_hd_temp_sd - Country_B_distance/Country_B_hd_temp_sd)^2
    # )
    ) %>% 
    # Ten years average of results
    group_by(taxon_key,ensemble,neighbour_id) %>% 
    mutate(RMean = rollmean(x = trans_index, 
                            10, 
                            align = "right", 
                            fill = trans_index)
    ) %>% 
    filter(year > 1960) %>%
    select(taxon_key,neighbour_id,year,ensemble,trans_index,RMean)

  # Average of 10 years mean of ensemble members results
  mean_sd_index <- Trans_index %>% 
    group_by(taxon_key,year,neighbour_id) %>% 
    summarise_at(vars("RMean"),
                 c(trend=mean,
                   ensemble_sd=sd,
                   n_ens = ~n())
    )

  # Estimate noice from ensemble members
  historic_sd <- Trans_index %>% 
    filter(year < Hist_y) %>% 
    group_by(taxon_key) %>% 
    summarise_at(vars("RMean"),
                 c(hist_mean=mean,
                   noise=sd,
                   n_ens = ~n())
    )

@@

  # If no historic data then emergin year is the first year of data
  if(nrow(historic_sd) == 0){

    Final_Result <- mean_sd_index %>%
      group_by(taxon_key,neighbour_id,n_ens) %>%
      summarise(emerging_yr= min(year)) %>%
      left_join(EEZ_ids,
                by = "neighbour_id") %>%
      ungroup() %>%
      select(taxon_key,eez_name,eez_neighbour,n_ens,emerging_yr)

  return(Final_Result)

  # Identifies the year of emergence (e.g. trend > noise)
  Final_Result <- mean_sd_index %>%
    filter(year > Hist_y,
           trend > (historic_sd$hist_mean+historic_sd$noise) |
             trend < (historic_sd$hist_mean-historic_sd$noise)
    ) %>%
    group_by(taxon_key,neighbour_id,n_ens) %>%
    summarise(emerging_yr= min(year)) %>%
    left_join(EEZ_ids,
              by = "neighbour_id") %>%
    ungroup() %>%
    select(taxon_key,eez_name,eez_neighbour,n_ens,emerging_yr)

  return(Final_Result)
  }
}

@@

GetTransIndex <- function(taxon_key,
                          year,
                          Hist_y,
                          data_type = "Catch",
                          path,
                          ensemble_list,
                          Neighbours,
                          eez_centroid){
  t_start <- Sys.time()
  print("Starting Step One")
    # First step
    # Call Get centroids fucntion for centroids data per ensemble
    Step_one <- bind_rows(
      lapply(ensemble_list,
             GetCentroids,
             taxon_key = taxon_key,
             year = year_df$year,
             data_type = data_type,
             Neighbours = Neighbours,
             eez_centroid = eez_centroid
             )
    )

@@

    print("Starting Step Two")

@@

    name <- paste(my_path("R",extra_path = "/Emergence/"),taxon_key,"_emergence.txt",sep="")

    write_csv(Step_two,
              name)

    time_dif <- Sys.time()-t_start

    message <- paste("Analysis done for", taxon_key, "in",time_dif, "minutes")
    return(message)
  }

  # GetTransIndex(
  #   taxon_key = 600076,
  #   ensemble_list = ensemble_list,
  #   year = year_df$year,
  #   Hist_y = 2000,
  #   data_type = "Catch",
  #   Neighbours = Neighbours,
  #   eez_centroid = eez_centroid
  # )

@@

# Already modeled species
spp_done <- list.files(my_path("R", extra_path = "Emergence"))

spp_done <- gsub("\\_.*","",spp_done)

spp_reman <- spp_list[!spp_list %in% spp_done]

@@

####_______________________ ####
# Estimating year of eergence
####_______________________ ####
n_cores <- 24
t_start <- Sys.time()

    Emerging_result <- bind_rows(
      lapply(spp_reman[63],
               GetTransIndex,
               year = year_df$year,
               Hist_y = Historic_year,
               data_type = "Catch",
               ensemble_list = ensemble_list,
               Neighbours = Neighbours,
               eez_centroid = eez_centroid
      )
    )

library(doParallel)
library(doSNOW)

cl <- makeCluster(12)
registerDoSNOW(cl)

# To show processing times
repetitions <- length(spp_reman)
pb <- txtProgressBar(max = repetitions, style = 3)

progress <- function(n) setTxtProgressBar(pb, n)
opts <- list(progress = progress)

results <- foreach(i=1:repetitions,
                    .combine = rbind,
                    .packages = c("MyFunctions",packages),
                    .options.snow = opts,
                   .verbose = T) %dopar% 
  GetTransIndex(
    taxon_key =spp_reman[i],
    ensemble_list = ensemble_list,
    year = year_df$year,
    Hist_y = 2000,
    data_type = "Catch",
    Neighbours = Neighbours,
    eez_centroid = eez_centroid
  )

close(pb)
stopCluster(cl)
